---
title: "Statistical Concepts"
author: "Lovette"
date: "2/5/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Loading Required Library
```{r}
library(tidyverse)
library(ggplot2)
```

Reading the dataset to R
```{r}
setwd("C:/Users/u/Documents/NLP work")
stroke_data <- read_csv("healthcare-dataset-stroke-data.csv")
```

Exploring the dataset
```{r}
glimpse(stroke_data)
```

```{r}
str(stroke_data)
```

Droping the 'id' column
```{r}

stroke_data <- stroke_data %>%
  select(-c(id))

```


Converting datatypes
```{r}
#character to factor
stroke_data[sapply(stroke_data, is.character)] <- lapply(stroke_data[sapply(stroke_data,
                                              is.character)], as.factor)
stroke_data$bmi <- as.numeric(as.character(stroke_data$bmi))
stroke_data$hypertension <- as.factor(stroke_data$hypertension)
stroke_data$heart_disease <- as.factor(stroke_data$heart_disease)
stroke_data$stroke <- as.factor(stroke_data$stroke)


```

```{r}
glimpse(stroke_data)
```

```{r}
str(stroke_data)
```


Checking for missing values
```{r}
colSums(is.na(stroke_data))
```

Removing missing values
```{r}
stroke_data <- stroke_data %>% na.omit()
```



Statistical summaries of dataset

```{r}
summary(stroke_data)
```


Exploring the distribution of the data set

```{r}

pass <- ggplot(stroke_data, aes(x=age)) + 
  geom_density()

# Add mean line
pass+ geom_vline(aes(xintercept=mean(age)),
            color="blue", linetype="dashed", size=1)
```

```{r}

pass <- ggplot(stroke_data, aes(x=avg_glucose_level)) + 
  geom_density()

# Add mean line
pass+ geom_vline(aes(xintercept=mean(avg_glucose_level)),
            color="blue", linetype="dashed", size=1)
```
The 'age' feature though slightly skewed can be assumed to have a normal distribution however, the 'avg_glucose_level' feature is left skewed.



Logistic Regression

Logistic regression is being used here because 'stroke' is encoded in binary format as it has on only 2 'factor' levels.

```{r}
lr_age1 <- glm(stroke ~ age, data = stroke_data, family=binomial)
summary(lr_age1)
```
The age is very significantly in predicting stroke and the difference between Null deviance and Residual deviance is reasonable telling us that the model is a good fit. 



```{r}
lr_bmi <- glm(stroke ~ bmi, data = stroke_data, family=binomial)
summary(lr_bmi)
```
The 'bmi' has a low significance in predicting 'stroke' and the difference between Null deviance and Residual deviance is also very small which signifies that this model is not a good fit.


```{r}
lr_gl <- glm(stroke ~ avg_glucose_level, data = stroke_data, family=binomial)
summary(lr_gl)
```
The 'avg_glucose_level' has a significance in predicting 'stroke' however, the difference between Null deviance and Residual deviance is low which signifies that this model is not a good fit.



Chi Squared Test
Chi Squraed test is being used as 'stroke' feature is categorical.

Reducing the data to frequencies for the categorical variables
```{r}
stroke_data_ct <- stroke_data %>% count(stroke, hypertension, sort = TRUE)
stroke_data_ct
```


Null hypothesis: independence ie having stroke is independent of the having 'hypertension' of the car.

Creating the matrix
```{r}
stroke_ct <- matrix(c(4309, 391, 149, 60), byrow = 2, nrow = 2, ncol = 2)
```

Applying the Chi test
```{r}
chisq.test(stroke_ct)
```
The p-value here is low (p-value < 0.05) therefore, we reject null hypothesis.


Chi Squared Test
Chi Squraed test is being used as 'stroke' feature is categorical.

Reducing the data to frequencies for the categorical variables
```{r}
stroke_data_ct <- stroke_data %>% count(stroke, heart_disease, sort = TRUE)
stroke_data_ct
```


Null hypothesis: independence ie having stroke is independent of the having 'heart_disease' of the car.

Creating the matrix
```{r}
stroke_ct <- matrix(c(4497, 203, 169, 40), byrow = 2, nrow = 2, ncol = 2)
```

Applying the Chi test
```{r}
chisq.test(stroke_ct)
```
The p-value here is low (p-value < 0.05) therefore, we reject null hypothesis.




T-test
Null Hypothesis: Mean is equal to 42.87

```{r}
t.test(stroke_data$age, mu = 42.87)
```
Since the p-value is greater than 0.05 we cannot reject the null hypothesis 


Analysis of Variance (ANOVA)

Null Hypothesis: There is no difference in average 'age' of the different 'smoking_status' categories.

One way ANOVA
```{r}
age_ss <- aov(age ~ smoking_status, data = stroke_data)
summary(age_ss)
```
Here we can obviously reject the null hypothesis as the p-value is very low.

Two way ANOVA
```{r}
age_hh <- aov(age ~ hypertension + heart_disease, data = stroke_data)
summary(age_hh)
```






Speraman's Correlation Test

Null Hypothesis: There is some correlation between 'avg_glucose_level' and the 'bmi'

```{r}
cor.test(stroke_data$avg_glucose_level, stroke_data$bmi, method = 'spearman', exact = FALSE)
```
From the correlation value of 0.1143, we can say that there is a positive correlation however, it is not a very strong correlation.


















